Positivity-Preserving Adaptive Runge--Kutta Methods

### 2020-05-04 nusslein_vsrp/latex/overview.tex
Remove reference to removed section.
Fix a few mistakes.

Our means to ensure positivity can be interpreted as a projection
approach, where the numerical solution is adapted to satisfy the
positivity constraint at the end of each step. In contrast to
simple orthogonal projection, which has also been proposed to deal
with positivity constraints \cite{shampine1986conservation},
our approach preserves all linear invariants of the given ODE. These
invariants can be very important. e.g.\ the total mass for a transport
problem or in reaction systems. Preserve of linear
invariants has been shown to be an important advantage of RRK methods
over orthogonal projection methods
\cite{ranocha2020relaxationHamiltonian}. Of course, it is also
possible to enforce the preservation of linear invariants in projection
methods, but the invariants have to be known explicitly
\cite{sandu2001positive}.

The paper unfolds as follows. In Section~\ref{sec:main_idea} the main idea is
explained. Section~\ref{sec:LP} contains the formulation of the linear program
for selection of the weights at each step.
Section~\ref{sec:integration} describes how the new approach can be used with
different RKMs, how it can be combined adaptive error control, and how the
region of absolute stability can be approximated.
Section~\ref{sec:imple} contains some details regarding efficient implementation 
of the algorithm.
In Section~\ref{sec:Numeric_Results} numerical results are given for multiple test problems.
A conclusion is given in Section~\ref{sec:conclusion}.

@@

The difference $\|\bt-b\|$ is also plotted in Figure\,\ref{fig:exampleI}.
No modification of the weights is required for $t<1.905$. At $t=1.905$ the
weights are first adapted to ensure the positivity of the solution. For
$t>2.63$ the original set of weights again lead to a positive solution,
and no further modification is necessary.

@@

Note that for any diagonally implicit method, by allowing the order to be reduced to one
we can guarantee the existence of a solution to the LP, since the backward Euler method
is unconditionally positive.  For explicit methods, reducing the order to one is guaranteed
to yield a solution of the LP only if the step size is small enough.

>>>>

Our means to ensure positivity can be interpreted as a projection
approach, where the numerical solution is adapted to satisfy the
positivity constraint at the end of each step. In contrast to
simple orthogonal projection, which has also been proposed to deal
with positivity constraints \cite{shampine1986conservation},
our approach preserves all linear invariants of the given ODE. These
invariants can be very important. e.g.\ the total mass for a transport
problem or in reaction systems. Preservation of linear
invariants has been shown to be an important advantage of RRK methods
over orthogonal projection methods
\cite{ranocha2020relaxationHamiltonian}. Of course, it is also
possible to enforce the preservation of linear invariants in projection
methods, but the invariants have to be known explicitly
\cite{sandu2001positive}.

The paper unfolds as follows. In Section~\ref{sec:main_idea} the main idea is
explained. Section~\ref{sec:LP} contains the formulation of the linear program
for selection of the weights at each step.
Section~\ref{sec:integration} describes how the new approach can be used with
different RKMs, how it can be combined with adaptive error control, and how the
region of absolute stability can be approximated.
In Section~\ref{sec:Numeric_Results} numerical results are given for multiple test problems.
A conclusion is given in Section~\ref{sec:conclusion}.

@@

The difference $\|\bt-b\|_1$ is also plotted in Figure\,\ref{fig:exampleI}.
No modification of the weights is required for $t<1.905$. At $t=1.905$ the
weights are first adapted to ensure the positivity of the solution. For
$t>2.63$ the original set of weights again lead to a positive solution,
and no further modification is necessary.

@@

Note that for many diagonally implicit methods, the first stage is a scaled backward Euler step.
For such methods, by allowing the order to be reduced to one
we can guarantee the existence of a solution to the LP, since the backward Euler method
is unconditionally positive.  For explicit methods, reducing the order to one is guaranteed
to yield a solution of the LP only if the step size is small enough.
